Specification and Verification of Stencil Computation Shape

### 2016-11-15 paper.tex

\section{Introduction}

@@

>>>>

\section{Introduction}\label{sec:intro}

@@

\section{Related work}\label{sec:related-work}

\citet{kamil2016verified} propose verified lifting to extract a mathematical
representation of low-level, and potentially optimised, stencils in Fortran
code. Problem domain wise their work is the closest to ours. However, since
their aim is to use the extracted predicate to pipe to the high-performance
compiler \textsf{Halide}~\citep{ragan2013halide}, they need to capture the full
meaning of the stencil computation.

This led them to develop \textsf{STNG}, a loop invariant and postcondition
finder through syntax-guided synthesis. Using this approach they restrict the
search space of postconditions to predicates of a certain form, one that is
compatible with \textsf{Halide}. They then look at concrete loop iterations and
form an hypothesis about the invariant, which is later confirmed or put back
into the system using a SMT solver.

CamFort differs from this work significantly here. The aim of CamFort is not a
full representation of the stencil, but enough information to capture the access
pattern. Hence, it does not require SMT solving or search of loop invariants. It
suffices to do comparatively simple data-flow analysis.

Due to the sheer amount of analysis involved, Kamil et al. reports that some
degenerate kernels take up to 17 hours to analyse and others require programmer
intervention for correct invariants to be inferred. Since CamFort mostly does
syntactic analysis, inference procedure always terminates quickly in all real
world examples and never requires programmer intervention. Downside of our
approach is that if optimisation causes stencil computation to be heavily
obfuscated, \textsf{STNG} would capture the access behaviour better.

More fundamentally, unlike CamFort, Kamil et al. does not provide a syntactic
representation of their postcondition, so there is no way of future-proofing the
code for future edits. Even if they were to provide a syntactic representation,
for complex stencils such as Navier-Stokes from \S~\ref{sec:intro}, it would be
as verbose as the code itself making it difficult for programmer to understand
the overall shape of the indexing behaviour.

\section{Conclusion}\label{sec:conclusion}
